---
title: "final_project"
author: "Storm Zierden!"
format: gfm
editor: source
execute: 
  echo: false
  warning: false
  message: false
  
---

```{r}
#| label: setup

library(tidyverse)
library("sf")
library("tigris")
library(tidyr)
```

```{r}
#| label: data
redlegged <-read_csv("redlegged-original/observations-714460.csv")

bullfrog <- read_csv("bullfrog-original/observations-714470.csv")
```
```{r}
#| label: shapefile

counties_sf <- counties(cb = TRUE, class = "sf", year = 2022)

```



```{r}
#| label: clean-redleg

cleanred <- redlegged |>
  select(c(id, observed_on, quality_grade, place_guess, latitude, longitude, species_guess, scientific_name, common_name)) |>
  drop_na() |>
filter(place_guess != "Ensenada, MX-BN, MX") |>
  separate_wider_delim(observed_on, delim = "-", names = c("year", "month", "day"))|>
st_as_sf(coords = c("longitude", "latitude"), crs = 4269) |>
   st_join(counties_sf, join = st_nearest_feature) |>
st_drop_geometry()
  

```
#https://copilot.microsoft.com/shares/KLbgNqjQ2ykxWrLaikS7E

```{r}
#| label: clean-bullfrog

cleanbull <- bullfrog |>
   select(c(id, observed_on, quality_grade, latitude, longitude, species_guess, scientific_name, common_name)) |>
  drop_na() |>
  separate_wider_delim(observed_on, delim = "-", names = c("year", "month", "day"))|>
st_as_sf(coords = c("longitude", "latitude"), crs = 4269) |>
  st_join(counties_sf, join = st_intersects) |>
  st_drop_geometry()

  
```

```{r}
#frog <-  cleanbull |>
  left_join(cleanred,
             by = c("year", "NAMESLAD"),
            relationship = "many-to-many")
```

```{r}
#| label: redfrog_viz
cleanred|>
  count(NAME, year) |>
  ggplot(aes(x = fct_inorder(NAME), y = n, fill = year)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "County Name", y = "Number of Frogs")
  
```
## Map does not look like how I want to, I want to seprate the red frog level by year and county in order to see its growth possible need make multiple and do facet but when i do it bunches it up a lot and does work how i want to
##I also want to combine the two data sets and see if the reason why some county have such low amounts is where bullfrogs have high population but when i join r isn't very happy with me
##i also need to clean up some of the data as it still have to much varabity to it.

